A Continual Planner that Switches Between Serial and Decision-Theoretic Planning

### 2010-11-17 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Belief revision.
 - Charles

Whereas until now we have considered a POMDP model factored in terms
of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations.  Where $\state, \state' \in
\states$, $a \in \actions$, from $\mu$ we have a state transition
function $\transProb(\state, \action, \state')$ giving the probability
of a transition from state $s$ to $s'$ if $a$ is applied. For any
$\state$ and $\action$ we have $\sum_{\state' \in \states}
\transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function.  Consequently there is a positive constant $c$
so that for all $\state \in \states$ and $\action \in \actions$,
$|\reward(\state, \action)| < c$.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

@@

Goal expressions in the problem description ---for example
$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
knows where the {\em cornflakes} are--- are treated as
control-knowledge by the planner. Here, CP is not allowed to start
executing a plan, unless in the finial stet the goal condition is
satisfied.

@@

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04}. PPDDL has been used to model
benchmarks in all 3 of the International Planning Competitions since
2004. A variation on PDDL for describing domains with stochastic
actions and uncertain starting configurations, PPDDL is a declarative
first-order (a.k.a., relational) that facilitates factored description
of domains and problem instances. There are straightforward
compilations from problems expressed compactly in PPDDL to
propositional representations amenable to state-of-the-art planning
procedures.  Because this language cannot model domains that feature
partial observability, for our setting we have developed DTPDDL, an
extension of PPDDL to partial observability.

>>>>

Whereas until now we have considered a POMDP model that is factored in
terms of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations.  Where $\state, \state' \in
\states$, $a \in \actions$, from $\mu$ we have a state transition
function $\transProb(\state, \action, \state')$ giving the probability
of a transition from state $s$ to $s'$ if $a$ is applied. For any
$\state$ and $\action$ we have $\sum_{\state' \in \states}
\transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function.  Consequently there is a positive constant $c$
so that for all $\state \in \states$ and $\action \in \actions$,
$|\reward(\state, \action)| < c$.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

Successive state estimation is available if we applying Bayes' rule.
Taking the current belief $\bstate$ as our {\em prior}, and supposing
action $\action$ is executed with perceptive outcome $\observ$, then
the probability that we are in $\state$ in the successive belief-state
$\bstate'$ is given by:

\begin{equation}\label{eq:revision}
\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
\end{equation}

\noindent where $\pp{Pr}(\observ | \action, \bstate)$ is the
normalising factor, that is, the probability of getting observation
$\observ$ given we execute $\action$ in $\bstate$.

@@

Goal expressions in the problem description ---for example
$(\pp{kval}~(\pp{location}~\pp{cornflakes}))$ expresses ``the robot
knows where the {\em cornflakes}--- are treated as control-knowledge
by the planner. Here, a plan is not considered valid unless in the
finial stet the goal condition is satisfied.

@@

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04}. PPDDL was used in all 3 of the
International Planning Competitions since 2004. A variation on PDDL
for describing domains with stochastic actions and uncertain starting
configurations, PPDDL is a declarative first-order (a.k.a.,
relational) that facilitates factored description of domains and
problem instances. There are straightforward compilations from
problems expressed compactly in PPDDL to propositional representations
amenable to state-of-the-art planning procedures.  Because this
language cannot model domains that feature partial observability, for
our setting we have developed DTPDDL, an extension of PPDDL to partial
observability.
